---
title: "BradyHW3"
format: html
editor: visual
---

# Homework 3 ST558 Group 2

## Task 1: Conceptual Questions

1.  To go from myfolder/homework/ to myfolder/MyData.csv would take ../MyData.csv since you have to go backwards.

2.  R projects are good for organizing, having the working directory be linked to the root of the project, gives good version control history with git and github.

3.  Git tracks our coding edits and changes. Github stores all this data and information from git itself.

4.  Tibbles keep their 2D structure after using '\[\]' instead of becoming a vector and also tibbles have special printing abilities like showing dimensions.

## Task 2: Reading Delimited Data

Reading in the glass data from URL in comma delimited file

```{r}
# reading in glass data
library(readxl)
library(tidyverse)

glass <- read_delim("https://www4.stat.ncsu.edu/~online/datasets/glass.data",
                   delim = ",",
                   col_names = c("Id", "RI", "Na", "Mg", "Al", "Si",
                                  "K", "Ca", "Ba", "Fe", "Type_of_Glass"))
glass
```

Reading in the yeast data from URL using read_table

```{r}
# reading in yeast data
yeast <- read_table("https://www4.stat.ncsu.edu/~online/datasets/yeast.data",
                    col_names = c("seq_name", "mcg", "gvh", "alm", "mit",
                                  "erl", "pox", "vac", "nuc", "class"))
yeast
```

## Task 3: Combining Excel and Delimited Data

Reading in white wine data with read_excel and renaming column

```{r}
# Reading in white wine excel data
white_wine <- read_excel("white-wine.xlsx")
white_wine

# Read the next sheet with variable names
white_names <- read_excel("white-wine.xlsx", sheet = 2)
white_names

# Renaming columns using the clean names from sheet 2
colnames(white_wine) <- white_names$Variables
white_wine

# Adding a column to indicate these wines are white
white_wine$type <- "white"
white_wine
```

Reading in red wine data from URL as semi colon delimited

```{r}
# Reading in red wine data (semicolon delimited) from URL
red_wine <- read_delim("https://www4.stat.ncsu.edu/~online/datasets/red-wine.csv",
                       delim = ";")
red_wine

# Renaming columns using the same clean names from sheet 2
colnames(red_wine) <- white_names$Variables
red_wine

# Adding a column to indicate these wines are red
red_wine$type <- "red"
red_wine
```

Combining the two data sets together

```{r}
# Combining red wine and white wine data sets into one
wine <- bind_rows(white_wine, red_wine)
wine
```

## Task 4: Writing a Function

Creating IQR Function, to calculate the interquartile range

```{r}
# Defining the IQR function
getIQR <- function(x, ...) {
  quants <- quantile(x, probs = c(0.25, 0.75), ...)
  iqr <- quants[2] - quants[1]     # 75th percentile minus 25th
  return(unname(iqr))              # remove the "75%" label
}
```

Use the IQR Function on the wine data set columns citric acid and fixed acidity

```{r}
# Applying IQR to fixed acidity and citric acid
getIQR(wine$fixed_acidity)
getIQR(wine$citric_acid)
```

Creating getCoefVar Function, to calculate coefficient of variation

```{r}
# Defining coefficient of variation
getCoefVar <- function(x, ...) {
  cv <- sd(x, ...) / abs(mean(x, ...))
  return(cv)
}
```

Use the getCoefVar Function on the wine data set columns citric acid and fixed acidity

```{r}
# Applying getCoefVar to fixed acidity and citric acid
getCoefVar(wine$fixed_acidity)
getCoefVar(wine$citric_acid)
```

Creating getSummaries, to return IQR and getCoefVar

```{r}
# Defining getSummaries
getSummaries <- function(x, ...) {
  iqr <- getIQR(x, ...)
  cv <- getCoefVar(x, ...)
  return(list(IQR = iqr, CoefVar = cv))
}
```

Use the getSummaries Function on the wine data set columns citric acid and fixed acidity

```{r}
# Applying getSummaries to fixed acidity and citric acid
getSummaries(wine$fixed_acidity)
getSummaries(wine$citric_acid)
```

Creating new infix function to replace cbind function, testing on the data frames and vector

```{r}
# Defining an infix function to replace cbind function
`%new_bind%` <- function(x, y) {
  cbind(x, y)   
}

# Setting up example data from HW3
df1 <- data.frame(unif = runif(10), norm = rnorm(10))
df2 <- data.frame(gamma = rgamma(10, shape = 1), beta = rbeta(10, shape1 = 5, shape2 = 1))
vec1 <- rnorm(10, mean = 10)

# Testing infix function on the data frames and vector
df1 %new_bind% df2
df1 %new_bind% df2 %new_bind% vec1
```
