---
title: "BradyHW3"
format: html
editor: visual
---

## Task 1: Conceptual Questions

1.  To go from myfolder/homework/ to myfolder/MyData.csv would take ../MyData.csv since you have to go backwards.

2.  R projects are good for organizing, having the working directory be linked to the root of the project, gives good version control history with git and github.

3.  Git tracks our coding edits and changes. Github stores all this data and information from git itself.

4.  Tibbles keep their 2D structure after using '\[\]' instead of becoming a vector and also tibbles have special printing abilities like showing dimensions.

## Task 2: Reading Delimited Data

Reading in the glass data from URL in comma delimited file

```{r}
# reading in glass data
library(tidyverse)

glass <- read_delim("https://www4.stat.ncsu.edu/~online/datasets/glass.data",
                   delim = ",",
                   col_names = c("Id", "RI", "Na", "Mg", "Al", "Si",
                                  "K", "Ca", "Ba", "Fe", "Type_of_Glass"))
glass
```

Reading in the yeast data from URL using read_table
```{r}
# reading in yeast data
yeast <- read_table("https://www4.stat.ncsu.edu/~online/datasets/yeast.data",
                    col_names = c("seq_name", "mcg", "gvh", "alm", "mit",
                                  "erl", "pox", "vac", "nuc", "class"))
yeast
```

## Task 3: Combining Excel and Deliminted Data

