---
title: "LoganHW3"
format: html
editor: visual
---

```{r}
#| include: FALSE
library(tidyverse)
library(readxl)
```


## Task 1

1)  We would use the path "../MyData.csv"

2)  It makes things easy to share, gives better integration with version control and git, and makes pathing easier.

3)  Github is a place where Git information is stored. Git is software that tracks changes in code.

4)  Tibbles are a specialized DF that prints differently and doesn't coerce a column to a vector when using [ notation.

## Task 2

### Reading in glass data with read_csv (URL)

```{r}
glassData <- read_csv("https://www4.stat.ncsu.edu/online/datasets/glass.data",
  col_names = FALSE,
  show_col_types = FALSE) 

names(glassData) <- c("Id", "RI", "Na", "Mg", "Al", "Si" , "K", "Ca", "Ba", "Fe", "Type of Glass")

# Could change ID and Type into chr if needed
glassData
```

### Reading in yeast data with read_table (URL)

```{r}
yeastData <- read_table("https://www4.stat.ncsu.edu/online/datasets/yeast.data",
  col_names = FALSE,
  show_col_types = FALSE) 

names(yeastData) <- c("seq_nameA", "mcg", "gvh", "alm", "mit", "erl" , "pox", "vac", "nuc", "class")

yeastData
```

## Task 3

### Reading in white wine data with read_excel

```{r}
# Reading in the data
wineWhiteData <- read_excel("white-wine.xlsx")

# Reading in names for the data from sheet 2
wineWhiteDataNames <- read_excel("white-wine.xlsx",
  sheet = excel_sheets("white-wine.xlsx")[2]
)

# Renaming
colnames(wineWhiteData) <- wineWhiteDataNames[[1]]

# Adding col to indicate wine type
wineWhiteData <- wineWhiteData %>% mutate(wine_type = "white")
```

### Reading in red wine data with read_delim

```{r}
# Reading in the data
wineRedData <- read_delim("red-wine.csv",
                          delim = ";",
                          col_names = TRUE,
                          show_col_types = FALSE)

# Renaming, has same names as white
colnames(wineRedData) <- wineWhiteDataNames[[1]]

# Adding col to indicate wine type also making sure everything is double for later operations
wineRedData <- wineRedData %>% mutate(across(wineWhiteDataNames[[1]], as.double)) %>% 
  mutate(wine_type = "red") 
```

### Combining the wine data

```{r}
wineData <- bind_rows(wineWhiteData, wineRedData)
```


## Task 4

### Creating function getIQR, uses quantile and does not have a name 

```{r}
getIQR <- function(x) {
  iqr <- quantile(x,.75) - quantile(x,.25)
  names(iqr) <- NULL
  return (iqr)
}
```

### getIQR output for columns fixed acidity and citric acid

```{r}
getIQR(wineData$fixed_acidity)
getIQR(wineData$citric_acid)
```

### Creating function getCoefVar

```{r}
getCoefVar <- function(x) {
  return (sd(x) / abs(mean(x)))
}
```

### getCoefVar output for columns fixed acidity and citric acid

```{r}
getCoefVar(wineData$fixed_acidity)
getCoefVar(wineData$citric_acid)
```
### Creating function getSummaries

```{r}
getSummaries <- function(x) {
  return (list(IQR = getIQR(x), CoefVar = getCoefVar(x)))
}
```

### getSummaries output for columns fixed acidity and citric acid

```{r}
getSummaries(wineData$fixed_acidity)
getSummaries(wineData$citric_acid)
```

### Infix function for cbind

```{r}
`%cb%` <- function(...) {
  return (cbind(...))
}
```

### Running infix cbind function

```{r}
# Example frames
df1 <- data.frame(unif = runif(10), norm = rnorm(10))
df2 <- data.frame(gamma = rgamma(10, shape = 1), beta = rbeta(10, shape1 = 5, shape2 = 1))
vec1 <- rnorm(10, mean = 10)

df1 %cb% df2
df1 %cb% df2 %cb% vec1
```
