---
title: "Mason HW3"
format: html
editor: visual
---

```{r}
#| include: FALSE

library(tidyverse)
library(readxl)
```

## Task 1: Conceptual Questions

1.  If our working directory is `myfolder/homework/` the relative path we would use to specify the file located at `myfolder/Mydata.csv` would be `../MyData.csv`.

2.  The major benefits of using R projects is that they integrate directly with Git/Github, the working directory being set to the project's root folder, and organization of files.

3.  `Git` is a version control software which keeps track of all files in an associated folder, including any changes made. `Github` is an online hosting service for Git-based projects, allowing you to have a remote file repository tracked by `Git`.

4.  The two main differences between a `tibble` and a `data frame` is that (1) tibbles do not coerce down to a vector when you subset to only one column using `[` and (2) they print cleaner than a data frame.

## Task 2: Reading Delimited Data


###Reading in glass data from the url https://www4.stat.ncsu.edu/online/datasets/glass.data

```{r}
glass_data <- read_csv("https://www4.stat.ncsu.edu/online/datasets/glass.data", col_names = F, show_col_types = F)

names(glass_data) <- c("Id", "RI", "Na", "Mg", "Al", "Si", "K", "Ca",
                       "Ba", "Fe", "Type of Glass")

glass_data
```
