---
title: "Mason HW3"
format: html
editor: visual
---

```{r}
#| include: FALSE

library(tidyverse)
library(readxl)
```

## Task 1: Conceptual Questions

1.  If our working directory is `myfolder/homework/` the relative path we would use to specify the file located at `myfolder/Mydata.csv` would be `../MyData.csv`.

2.  The major benefits of using R projects is that they integrate directly with Git/Github, the working directory being set to the project's root folder, and organization of files.

3.  `Git` is a version control software which keeps track of all files in an associated folder, including any changes made. `Github` is an online hosting service for Git-based projects, allowing you to have a remote file repository tracked by `Git`.

4.  The two main differences between a `tibble` and a `data frame` is that (1) tibbles do not coerce down to a vector when you subset to only one column using `[` and (2) they print cleaner than a data frame.

## Task 2: Reading Delimited Data

###Reading in glass data from the url https://www4.stat.ncsu.edu/online/datasets/glass.data

```{r}
#Reading in the data
glass_data <- read_csv("https://www4.stat.ncsu.edu/online/datasets/glass.data", col_names = F, show_col_types = F)

#Setting variable names based on information from the homework
names(glass_data) <- c("Id", "RI", "Na", "Mg", "Al", "Si", "K", "Ca",
                       "Ba", "Fe", "Type of Glass")

#Printing tibble
glass_data
```

### Reading in the yeast data now from the url https://www4.stat.ncsu.edu/online/datasets/yeast.data

```{r}
#Reading in the data
yeast_data <- read_table("https://www4.stat.ncsu.edu/online/datasets/yeast.data", col_names = F, show_col_types = F)

#Setting variable names based on information from the homework
names(yeast_data) <- c("Seq Name", "MCG", "GVH", "ALM", "MIT", "ERL",
                       "POX", "VAC", "NUC", "Class")

#Printing the tibble
yeast_data
```

## Task 3: Combining Excel and Delimited Data
### Reading in the white wine data

```{r}
#Reading in data from the first sheet
white_data <- read_excel("white-wine.xlsx", sheet = excel_sheets("white-wine.xlsx")[1])

#Printing the tibble
white_data
```

Reading in the second sheet to get the variable names
```{r}
#Reading in the second sheet of the excel data
white_names <- read_excel("white-wine.xlsx", sheet = excel_sheets("white-wine.xlsx")[2])

#Setting the variable names to the values from the white_names tibble
colnames(white_data) <- white_names[[1]]
```

Adding a column to the white wines dataset indicating that the wine is white

```{r}
white_data <- white_data %>% mutate(wine_type = "white")
```

###Reading in the red wine data 
```{r}
#Using delim = ";" to tell R how to separate the values
red_data <- read_delim("red-wine.csv", delim = ";", col_names = T,
                     show_col_types = F)

#Setting the variable names to the values from the white_names tibble
colnames(red_data) <- white_names[[1]]

#Creating a variable that indicates that the wine is red
red_data$wine_type = "red"
```

Now I am merging both datasets
```{r}
combined_wines <- bind_rows(white_data, red_data)

combined_wines
```

## Task 4: Writing a Function

Creating the `getIQR()` function that takes in a vector of numbers and returns the interquartile range.

```{r}
getIQR <- function(x){
  iqr <- quantile(x, probs = .75) - quantile(x, probs = .25)
  names(iqr) <- NULL
  return (iqr)
}
```

Using our `getIQR()` function to return the IQR of the `fixed_acidity` and `citric_acid` columns from the combined wines data set.

```{r}
getIQR(combined_wines$fixed_acidity)
getIQR(combined_wines$citric_acid)
```

Creating the `getCoefVar()` function to return the sample coefficient of variation by dividing the sample standard deviation by the sample mean

```{r}
getCoefVar <- function(x){
  CoefVar <- sd(x) / mean(x)
  names(CoefVar) <- NULL
  return(CoefVar)
}
```

Using our `getCoefVar()` function to return the sample variance of the `fixed_acidity` and `citric_acid` columns from the combined wines data set.

```{r}
getCoefVar(combined_wines$fixed_acidity)
getCoefVar(combined_wines$citric_acid)
```

Creating the `getSummaries()` function to return the IQR and the sample coefficient of variation by calling our previous functions

```{r}
getSummaries <- function(x){
  return(list(IQR = getIQR(x), CoefVar = getCoefVar(x)))
}
```

Using our `getSumamries()` function to return the IQR and the sample variance of the `fixed_acidity` and `citric_acid` columns from the combined wines data set.

```{r}
getSummaries(combined_wines$fixed_acidity)
getSummaries(combined_wines$citric_acid)
```

Creating our own infix function to replace the `cbind()` function

```{r}
`%new_bind%` <- function(...){
  return(cbind(...))
}
```

Reading in the data frames given to us in the homework and applying our infix function to check that it works

```{r}

```