---
title: "Ryan's HW3"
format: html
---

## Task 1: Conceptual Questions

1. Given a working directory of $\texttt{myfolder/homework/}$, the relative path we would use to specify a file at $\texttt{myfolder/MyData/}$ would be $\texttt{`../MyData.csv`}$. 

2. The major benefits of using R projects are organization of files, integration with version control, and relative path consistency due to the working directory being set to the project root.

3. Git is a system for version control. It is used to track changes and save prior versions of work. Github is a cloud platform used primarily for software development. It hosts Git repositories and allows for a collaborative environment.  

4. The two main differences between a $\texttt{tibble}$ and a $\texttt{data.frame}$ are: 1) Tibbles preserve their 2D structure after using '[', and 2) tibbles allow for special printing.

## Task 2: Reading Delimited Data

We are reading in the glass data from its URL, which is comma-delimited with 11 name-less columns. 
```{r}
library(tidyverse)

# Read in glass data
glass_data <- read_csv(
  "https://www4.stat.ncsu.edu/~online/datasets/glass.data",
  col_names = c("Id", "RI", "Na", "Mg", "Al", "Si", "K", "Ca", "Ba", "Fe", "Type")
)

# Print tibble
glass_data
```

Similarly as above, we are to read in data from a URL and print the tibble. This time it's with the yeast data, which has a fixed width field and 10 columns.
```{r}
# Read in yeast data
yeast_data <- read_fwf(
  "https://www4.stat.ncsu.edu/~online/datasets/yeast.data",
  fwf_widths(
    widths = c(12, 6, 6, 6, 6, 6, 6, 6, 6, 3),               # Character widths for each column
    col_names = c("seq_name", "mcg", "gvh", "alm", "mit",
                  "erl", "pox", "vac", "nuc", "class")
  )
)

# Print tibble
yeast_data
```
## Task 3: Combining Excel and Delimited Data

Now we will read in the white wine data set, which is an Excel data set with two sheet. The data set has been stored at the level of this .qmd file but not committed. 
```{r}
library(readxl)

# Read in white wine data
white_wine <- read_excel("white-wine.xlsx")

# Print tibble
white_wine
```

Renaming variables with those labels from the second sheet.
```{r}
# Specifying the second sheet
white_wine_names <- read_excel("white-wine.xlsx", sheet = 2)

# Renaming columns
colnames(white_wine) <- white_wine_names$Variables

# Print tibble
white_wine
```

Adding a column to indicate the wines are white.
```{r}
# Column add
white_wine$type <- "white"

# Print tibble
white_wine
```

Next, we will read in the red wine data set, which is semicolon-delimited. Afterwards, we will perform the same modifications as we did for white wine.
```{r}
# Read in red wine data
red_wine <- read_delim("red-wine.csv", delim = ";")

# Renaming columns
colnames(red_wine) <- white_wine_names$Variables

# Column add
red_wine$type <- "red"

# Print tibble
red_wine
```

Finally, we shall merge the red and white wine data sets we've created.
```{r}
# Combine data sets
wine_data <- dplyr::bind_rows(white_wine, red_wine)

# Print tibble
wine_data
```

## Task 4: Writing a Function

We create the getIQR function that determines the interquartile range for a provided vector of numbers.
```{r}
# Define function
getIQR <- function(x, ...) {                             # ... allows for other arguments to be passed
  quartiles <- quantile(x, probs = c(0.25, 0.75), ...)   # 25th and 75th percentiles
  iqr <- quartiles[[2]] - quartiles[[1]]                 # Removing name on vector and compute IQR
  iqr
}
```

Applying the getIQR function to our merged wine data set above (i.e. wine_data), specifically the fixed_acidity and citric_acid columns. 
```{r}
# IQR for the combined wine data set
getIQR(wine_data$fixed_acidity)
getIQR(wine_data$citric_acid)
```

Creating the getCoefVar function to find the sample coefficient of variation. 
```{r}
# Define function
getCoefVar <- function(x, ...) {
  cv <- sd(x, ...) / abs(mean(x, ...))  # Sample coeff of variation equation
  cv
}
```

Applying the getCoefVar function to our merged wine data set above (i.e. wine_data), specifically the fixed_acidity and citric_acid columns. 
```{r}
# Sample coefficient of variation for the combined wine data set
getCoefVar(wine_data$fixed_acidity)
getCoefVar(wine_data$citric_acid)
```

Creating the getCoefVar function to return the IQR and coefficient of variation. 
```{r}
# Define function
getSummaries <- function(x) {
  IQR_val <- getIQR(x)
  CV_val <- getCoefVar(x)
  return(list(IQR = IQR_val, CoefVar = CV_val)) # Needed to ensure both values are returned
}
```

Applying the getSummaries function to our merged wine data set above (i.e. wine_data), specifically the fixed_acidity and citric_acid columns. 
```{r}
# IQR and coefficient of variation for the combined wine data set
getSummaries(wine_data$fixed_acidity)
getSummaries(wine_data$citric_acid)

```

Create a new infix function to replace cbind.
```{r}
# Define function
`%new_bind%` <- function(x, y) {
  cbind(x, y)               # Taking advantage of cbind's built-in properties
}

# Example data frames
df1 <- data.frame(unif = runif(10), norm = rnorm(10))
df2 <- data.frame(gamma = rgamma(10, shape = 1), beta = rbeta(10, shape1 = 5, shape2 = 1))
vec1 <- rnorm(10, mean = 10)

# Testing function
df1 %new_bind% df2
```

```{r}
# Testing function
df1 %new_bind% df2 %new_bind% vec1  # Note that y = vec1
```

