---
title: "JingyiZuo_homework2 "
format: html
---

# Task 1 
1. path: /myfolder 
2. it helps us to manage different docments for one project, complete version control with Git, and easy to collaborate with others. 
3. Git is a tool to help us to do version control.Github is a platform we could use to manage project, collaborate with others, and show opensource codes on the internet. 
4. Tibble is a special version of a dataframe. Tibble's classes include dataframe, but also include tbl_df and tbl. Tidyverse uses tibble for both of input and output. 

# Task 2 
1. glass data
```{r glass data}
library(tidyverse)
glass_tbl <- read_delim("https://www4.stat.ncsu.edu/~online/datasets/glass.data", delim = ",", col_names = c("Id","RI","Na","Mg","Al","Si","K","Ca","Ba","Fe","Type of Class"))

glass_tbl 


```



2. Yeast data
```{r Yeast data}
library(tidyverse)
yeast_tbl <- read_delim("https://www4.stat.ncsu.edu/~online/datasets/yeast.data", delim = " ", col_names = c("seq_name","mcg","gvh","alm","mit","erl","pox","vac","nuc","class"))
yeast_tbl
```

# Task 3 
```{r combine excel and delimited data}
# read 1st sheet 
library(readxl)
whitewine <- read_excel("white-wine.xlsx",
                        sheet = excel_sheets("white-wine.xlsx")[1])
whitewine

# read 2st sheet
whitewine_2 <- read_excel("white-wine.xlsx",
                        sheet = excel_sheets("white-wine.xlsx")[2])
colnames(whitewine) <- whitewine_2$Variables[1:13]
whitewine

# add type column and assign "white" as type 
whitewine$type <- "white"

# read another excel 
library(readr)
redwine <- read_delim("https://www4.stat.ncsu.edu/~online/datasets/red-wine.csv", delim = ";")
colnames(redwine) <- whitewine_2$Variables[1:13]
redwine$type <- "red"

# combine two excels
allwine <- dplyr::bind_rows(whitewine, redwine)
```
# task 4 
```{r write a function }
# define the function
getIQR <- function(x) {a <- quantile(x, 0.25, na.rm = TRUE) 
b<- quantile(x,0.75, na.rm = TRUE)
c <- unname(b-a) 
c}

# 9 
getIQR(allwine$fixed_acidity)
getIQR(allwine$citric_acid)

# 10 define anther function
getCoefVar <- function (x) {mean <- mean(x, na.rm = TRUE)
sd <- sd(x, na.rm= TRUE) 
sd/mean} 

getCoefVar(allwine$fixed_acidity)
getCoefVar(allwine$citric_acid)

# 12 define summary function
getSummaries <- function(x) {a <- getIQR(x)
  b <- getCoefVar(x)
  list(IQR = a, coefVae = b)
}

getSummaries(allwine$fixed_acidity)
getSummaries(allwine$citric_acid)

# 14 define infix function
infix <- function(df1 , df2, vec1 = NULL) { if (is.null(vec1)) cbind(df1, df2)
else cbind(df1,df2,vec1)} 
df1 <- data.frame(unif = runif(10), norm = rnorm(10))
df2 <- data.frame(gamma = rgamma(10, shape = 1), beta = rbeta(10, shape1 = 5, shape2 = 1))
vec1 <- rnorm(10, mean = 10)
infix(df1, df2)
infix(df1, df2, vec1)
```