---
title: "BradyHW2"
format: html
editor: visual
---

# Task 1: Basic Vector Practice

I created pre and post treatment vectors

```{r}
# Pre Treatment vector 20 observations
pre <- c(130, 128, 116, 124, 133, 134, 118, 126, 114, 127, 141, 138, 128, 140, 137, 131, 120, 128, 139, 135)
# Post Treatment vector 20 observations
post <- c(114, 98, 113, 99, 107, 116, 113, 111, 119, 117, 101, 119, 130, 122, 106, 106, 124, 102, 117, 113)
```

I assigned the pre and post vectors names of pre and post

```{r}
# Assigning names to vector elements
names(pre) <- paste("Subject", 1:20, sep = "_")
names(post) <- paste("Subject", 1:20, sep = "_")
```

I calculated the difference between the before and after and called this difference diff

```{r}
# Difference in blood pressure pre and post 
diff <- pre - post
diff
```

I calculated the average change in blood pressure of all patients

```{r}
# Average of the difference of bp
mean(diff)
```

I determine which patients blodd pressures decreased, which is a positive change

```{r}
# Patients with positvie change in bp list
which(diff>0)
```

I subset the diff vector to show only patients with positive bp change

```{r}
#keep only positive bp changes
pos_diff <- diff[diff>0]
pos_diff
```

I calculated average decrease in bp for only patients who bp decreased

```{r}
# Average decrease among patients who decreased
mean(pos_diff)
```

## Task 2: Basic Data Frame Practice

I creates data frame object with 4 columns
```{r}
# 4 new columns in Data Frame
df_bp <- data.frame(patient = names(pre),
                    pre_bp = pre,
                    post_bp = post,
                    diff_bp = diff)
```

I get only the patients with diff_bp is negative.

```{r}
# Only diff_bp negative rows
df_bp[df_bp$diff_bp<0, ]
```

I created a new column for if post_bp is under 120
```{r}
# Adding a column, read back TRUE if post_bp is under 120
df_bp$below_120 <- df_bp$post_bp<120
df_bp
```

I printed out the data frame
```{r}
#Print data frame using knitr
knitr::kable(df_bp)
```

