---
title: "BradyHW5"
format: html
editor: visual
---

# Homework 5 ST 558 Group 2

```{r}
#| include: false
library(tidyverse)
library(httr)
library(jsonlite)
```

## Task 1: Querying an API and a Function

Storing my API key and connecting to newsapi.org for recent news

```{r}
# My newsapi key
my_key <- "8d4baeba91e14e559c0393b7151d48d8"

# Grab recent news about weather, going back 30 days (free account limit)
from_date <- Sys.Date() - 30

news_result <- GET(paste0(
  "https://newsapi.org/v2/everything?",
  "q=weather&",
  "from=", from_date, "&",
  "language=en&",
  "pageSize=50&",
  "apiKey=", my_key
))
```

Parsing the result down to the article data frame

```{r}
# Parse the returned JSON and pull the articles into a tibble
parsed <- fromJSON(rawToChar(news_result$content))
articles <- as_tibble(parsed$articles)
articles
```

Writing a function so a user can query the API easily

```{r}
# Function to query newsapi: takes a search term, a from date, and an api key
query_news <- function(search_term, from_date, api_key) {
  result <- GET(paste0(
    "https://newsapi.org/v2/everything?",
    "q=", search_term, "&",
    "from=", from_date, "&",
    "language=en&",
    "pageSize=50&",
    "apiKey=", api_key
  ))
  parsed <- fromJSON(rawToChar(result$content))
  articles <- as_tibble(parsed$articles)
  return(articles)
}
```

Testing the function twice with different topics

```{r}
# Two test queries
query1 <- query_news("sports", as.character(Sys.Date() - 30), my_key)
query2 <- query_news("music", as.character(Sys.Date() - 30), my_key)

query1
query2
```

## Task 2: Summarizing Roller Coaster Data

Reading in the roller coaster data and checking how it's stored

```{r}
# Read in the roller coaster data
rc_data <- read_csv("https://www4.stat.ncsu.edu/~online/datasets/635USrollercoasters.csv")

# Check the structure
str(rc_data)
```

### Basic EDA

Converting the categorical variables to factors, then checking missing values

```{r}
# AgeGroup as an ordered factor (Older < Recent < Newest)
rc_data$AgeGroup <- factor(rc_data$AgeGroup,
                           levels = c("1:Older", "2:Recent", "3:Newest"),
                           ordered = TRUE)

# Other categorical variables to factor (backticks for names with spaces or symbols)
rc_data$Scale <- factor(rc_data$Scale)
rc_data$Type <- factor(rc_data$Type)
rc_data$Design <- factor(rc_data$Design)
rc_data$`Census Region` <- factor(rc_data$`Census Region`)
rc_data$`Inversions?` <- factor(rc_data$`Inversions?`)

# Check missing values in each column
colSums(is.na(rc_data))
```

A few columns have missing values: TopSpeed, MaxHeight, Drop, Length, Duration, and Make. These are mostly ride specification numbers, plus the make for some rides. None of the factor columns are missing.

### Categorical Variables

Making a one-way table on coaster Type

```{r}
# One-way table: Type
one_way <- table(rc_data$Type)
one_way
```

There are 530 steel coasters out of 605 total, so the large majority of coasters are steel rather than wood.

Making a two-way table of Type by Scale

```{r}
# Two-way table: Type by Scale
two_way <- table(rc_data$Type, rc_data$Scale)
two_way
```

Looking at where Type is Wood and Scale is Family, that tells us how many wood coasters are classified as family rides. There is only 1 wood coaster that is a kiddie ride. Since most of the coasters are steel, it makes since that there are more steel coasters of all Scales compared to Wood coasters.

Making a three-way table of Type by Scale by Census Region

```{r}
# Three-way table: Type by Scale by Census Region
three_way <- table(rc_data$Type, rc_data$Scale, rc_data$`Census Region`)
three_way
```

Picking the West region, Wood, Family a 0 there means there are no wooden family coasters in the West, unlike the other regions. Across all regions, there seems to be a close number of wooden thrill coasters of 5, 9, 11, and 11 as well as kiddie steel coasters with 8, 6, 6, and 7.

Making a conditional two-way table two different ways, conditioning on the South region

```{r}
# Method 1: filter to South first, then make the two-way table
south_data <- filter(rc_data, `Census Region` == "South")
table(south_data$Type, south_data$Scale)
```

```{r}
# Method 2: pull the South slice out of the three-way table
three_way[, , "South"]
```

Recreating the Type by Scale two-way table using dplyr

```{r}
# Group and count, then pivot wider to look like table() output
rc_data |>
  group_by(Type, Scale) |>
  summarize(count = n(), .groups = "drop") |>
  pivot_wider(names_from = Scale, values_from = count)
```

### Bar Graphs

Making a stacked and a side-by-side bar graph of Scale by Type

```{r}
# Turn the two-way table into a data frame for plotting
type_scale_df <- as.data.frame(two_way)

# Stacked bar graph
ggplot(type_scale_df, aes(x = Var1, y = Freq, fill = Var2)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(x = "Coaster Type", y = "Number of Coasters",
       title = "Roller Coaster Scale by Type", fill = "Scale")

# Side-by-side bar graph
ggplot(type_scale_df, aes(x = Var1, y = Freq, fill = Var2)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Coaster Type", y = "Number of Coasters",
       title = "Roller Coaster Scale by Type", fill = "Scale")
```
