---
title: "Mason HW5"
format: html
editor: visual
---

```{r}
library(httr)
library(tidyverse)
library(jsonlite)
library(ggplot2)
library(janitor)
```

## Task 1 - Querying an API and a Tidy-Style Function

Connecting to newsapi.org and grabbing news from the past 30 days about the NFL

```{r}
url <- "https://newsapi.org/v2/everything?q=nfl&from=2026-08-25&apiKey=654641465c3f469891708ce3b77fb2a0"

nflReturn <- GET(url)
parsedNFL <- fromJSON(rawToChar(nflReturn$content))
nflInfo <- as_tibble(parsedNFL)
```

Creating function to easily query this API

```{r}
#The function takes in 3 parameters from the user, the topic they are interested in, the time frame, and the users API key
queryAPI <- function(topic, period, key){
  url <- paste0("https://newsapi.org/v2/everything?q=", topic, "&from=",
                period, "&apiKey=", key)
  ret <- GET(url)
  parsed <- fromJSON(rawToChar(ret$content))
  return(as_tibble(parsed))
}
```

Testing our function works

```{r}
news_test_1 <- queryAPI("usa", "2026-08-25", "654641465c3f469891708ce3b77fb2a0")

news_test_2 <- queryAPI("apple", "2026-08-25", "654641465c3f469891708ce3b77fb2a0")
```

## Task 2 - Summarizing Roller Coaster Data

Reading in roller coaster data and converting necessary variables to factors
```{r}
coasters <- read_csv("635USrollercoasters.csv", show_col_types = F) %>%
  clean_names() %>%
  mutate(across(c("census_region","scale","type","design","inversions"),
                as.factor)) %>%
  mutate(age_group = factor(age_group, 
                            c("1:Older", "2:Recent", 
                                         "3:Newest"), 
                            c("Older", "Recent", "Newest")))
```

Checking for any missing values
```{r}
colSums(is.na(coasters))
```
We can see there are missing values for the following variables: top_speed, max_height, drop, length, duration, and make

Creating one-way contingency table
```{r}
oneway <- table(coasters$scale)
oneway
```
This one-way contingency table tells us that there are 269 roller coasters with an extreme scale

Creating two-way contingency table
```{r}
twoway <- table(coasters$scale, coasters$age_group)
twoway
```
This two-way contingency table tells us that out of the 269 extreme scale coasters, 17 are considered older

Creating three-way contingency table
```{r}
threeway <- table(coasters$scale, coasters$age_group, coasters$type)
threeway
```
There are 8 roller coasters that are constructed with steel, considered extreme, and are older

Creating conditional two-way table using three-way table from above

```{r}
threeway[, , "Steel"]
```
Creating conditional two-way table using `filter()`

```{r}
coasters %>%
  filter(type == "Steel") %>%
  tabyl(scale, age_group)
```

Creating two-way contingency table using `group_by()` and `summarize()`

```{r}
coasters %>%
  group_by(scale, age_group) %>%
  summarize(count = n()) %>%
  pivot_wider(names_from = age_group, values_from = count)
```

Creating stacked bar chart
```{r}
ggplot(data = coasters |> drop_na(scale, age_group), 
       aes(x = scale, fill = age_group)) +
  geom_bar() +
  labs(title = "Stacked Bar Chart of Coaster's Scale by \nAge of the Coaster", x = "Scale", 
       y = "# of Coasters", fill = "Age Group") +
  theme_minimal()
```

Creating side-by-side bar graph for roller coaster scale and age group

```{r}
ggplot(data = coasters |> drop_na(scale, age_group), 
       aes(x = scale, fill = age_group)) +
  geom_bar(position = "dodge") +
  labs(title = "Side-by-Side Graph of Coaster's Scale by \nAge of the Coaster", 
       x = "Scale", 
       y = "# of Coasters", fill = "Age Group") +
  theme_minimal()
```
