---
title: "ZUOHW5"
format: html
---

## Task1

```{r}
# use API to request data from NEWS API, no more than 50 news after 08/30 about happy in English.

library(httr)
library(jsonlite)
# Store the API as an environment variable and reference it when access data.
  Sys.setenv(API_KEY = "9e08bf550c3649eda8f016634bafce69")
 news <- GET("https://newsapi.org/v2/everything?q=happy&from=2026-08-30&language=en&pageSize=50",
      query = list(apiKey=Sys.getenv("API_KEY")))
 
### parse the return
 txt <- content(news, as = "text", encoding = "UTF-8")

result <- fromJSON(txt, flatten = TRUE)
result 

```

```{r}
## create a function

reqnews <- function(title, period, apikey){
  news <- GET("https://newsapi.org/v2/everything",            query = list(q= title, 
                                                                                  `from` = period, 
                                                                                  language ="en", 
                                                                                  pageSize = 50,
                                                                                  apiKey=apikey))
  txt <- content(news, as = "text", encoding = "UTF-8")
result <- fromJSON(txt, flatten = TRUE)
result
}

## try this function for the first time
reqnews("apple", "2026-09-03", Sys.getenv("API_KEY"))
## try this function for the second time 
reqnews("gender", "2026-09-21", Sys.getenv("API_KEY"))
```
### Task2 EDA
```{r}
# read data from csv 
library(tidyverse)
rocoaster <- read_csv("https://www4.stat.ncsu.edu/online/datasets/635USrollercoasters.csv")

# basic EDA
library(tidyverse)
dim(rocoaster)
names(rocoaster)
glimpse(rocoaster)
head(rocoaster)
summary(rocoaster)
str(rocoaster)
head(rocoaster,5)

# 
rocoaster <- rocoaster |>
  mutate (
    across(
      c(State, `Census Region`, AgeGroup, Scale, Type, Design, `Inversions?` ),
      as.factor
    )
  )


```



```{r}
#missing values examination
colSums(is.na(rocoaster))
rocoaster |> summarize(across(everything(), ~mean(is.na(.x))))


```
11.7% in Topseed missing (n = 74), 1.3% in MaxHight missing (n = 8), 20.16% in Drop missing (n = 128); 5.2% in Length missing (n = 33), 10.4% in Duration missing (n=66). 

## Task 3 Categorical analysis 

```{r}
table(rocoaster$AgeGroup)
table(rocoaster$AgeGroup, rocoaster$Scale)
table(rocoaster$AgeGroup, rocoaster$Scale, rocoaster$Type)


```
69 in one-way table means There are 69 roller coasters are older; "90" in two-way table means there are 90 thriller-type roller coasters are newest. "37" in three ways table means there are 37 family-type roller coaster made of stell are recently built. 


conditional two-way table
```{r}


ro2 <- rocoaster |>
  filter(Type == "Steel")

table(ro2$AgeGroup, ro2$Scale)

tab1 <- table(rocoaster$AgeGroup, rocoaster$Scale, rocoaster$Type)

tab2 <- tab1[,,"Steel"]
tab2


```

two-way contingency table and use pivot_wider 

```{r pivot_wider}
library(dplyr)

table2 <- rocoaster |>
  group_by (AgeGroup,Scale) |>
  summarize (n = n(), .groups = "drop") |>
  pivot_wider(
    names_from = AgeGroup,
    values_from = n,
    values_fill = 0)


table2 

```
stacked bar graph and side-by-side bae gragh 
```{r}
library(ggplot2)
# Stacked bar plot
ggplot(rocoaster, aes(x=Scale, fill = AgeGroup) ) + 
  geom_bar() +
  labs(
    title = "the relationship between rollercoaster's type and age",
    x = "age",
    y = "count",
    fill = "age"
  )

# Side-by-side 
ggplot(rocoaster, aes(x=Scale, fill = AgeGroup) ) +
  geom_bar(position = "dodge") + 
labs(
    title = "the relationship between rollercoaster's type and age",
    x = "age",
    y = "count",
    fill = "age"
  )
```
