---
title: "ZUOHW4"
format: html
---

## Task1

```{r}
# use API to request data from NEWS API, no more than 50 news after 08/30 about happy in English.

library(httr)
library(jsonlite)
# Store the API as an environment variable and reference it when access data.
  Sys.setenv(API_KEY = "9e08bf550c3649eda8f016634bafce69")
 news <- GET("https://newsapi.org/v2/everything?q=happy&from=2026-08-30&language=en&pageSize=50",
      query = list(apiKey=Sys.getenv("API_KEY")))
txt <- content(news, as = "text", encoding = "UTF-8")
result <- fromJSON(txt, flatten = TRUE)
result 

```
### Task2 EDA
```{r}
# read data from csv 
rocoaster <- read_delim("https://www4.stat.ncsu.edu/online/datasets/635USrollercoasters.csv", delim = ",")

# basic EDA
library(tidyverse)
dim(rocoaster)
names(rocoaster)
glimpse(rocoaster)
head(rocoaster)
summary(rocoaster)
str(rocoaster)
head(rocoaster,5)

# 
rocoaster <- rocoaster |>
  mutate (
    across(
      c(State, `Census Region`, AgeGroup, Scale, Type, Design, `Inversions?` ),
      as.factor
    )
  )


```



```{r}
#missing values examination
colSums(is.na(rocoaster))
rocoaster |> summarize(across(everything(), ~mean(is.na(.x))))


```
11.7% in Topseed missing (n = 74), 1.3% in MaxHight missing (n = 8), 20.16% in Drop missing (n = 128); 5.2% in Length missing (n = 33), 10.4% in Duration missing (n=66). 

## Task 3 Categorical analysis 

```{r}
table(rocoaster$AgeGroup)
table(rocoaster$AgeGroup, rocoaster$Scale)
table(rocoaster$AgeGroup, rocoaster$Scale, rocoaster$Type)


```
69 in one-way table means There are 69 rollercoaster are older; 90 in two-way table means there are 90 roller coaster are Newest but belong to thrill type. 37 in three ways table means there are 37 roller coaster made of stell are Recent built and belong to family type. 


```{r}
ro2 <- rocoaster |>
  filter(Type == "Steel")

table(ro2$AgeGroup, ro2$Scale)

tab1 <- table(rocoaster$AgeGroup, rocoaster$Scale, rocoaster$Type)

tab2 <- tab1[,,"Steel"]
tab2


```



```{r}
library(ggplot2)
# Stacked bar plot
ggplot(rocoaster, aes(x=Scale, fill = AgeGroup) ) + 
  geom_bar() +
  labs(
    title = "the relationship between rollercoaster's type and age",
    x = "age",
    y = "count",
    fill = "age"
  )

# Side-by-side 
ggplot(rocoaster, aes(x=Scale, fill = AgeGroup) ) +
  geom_bar(position = "dodge") + 
labs(
    title = "the relationship between rollercoaster's type and age",
    x = "age",
    y = "count",
    fill = "age"
  )
```
