---
title: "St558hw2"
format: html
editor: visual
---

## Task 1

### Vectors for data

```{r}
preTreat <- c(130, 128, 116, 124, 133, 134, 118, 126, 114, 127, 141, 138, 128, 140, 137, 131, 120, 128, 139, 135)

postTreat <- c(114, 98, 113, 99, 107, 116, 113, 111, 119, 117, 101, 119, 130, 122, 106, 106, 124, 102, 117, 113)

names(preTreat) <- paste("Subject", 1:20, sep = "_")

names(postTreat) <- paste("Subject", 1:20, sep = "_")
```

### Change in blood pressure

```{r}
changeTreat <- preTreat - postTreat

avgChange <- mean(changeTreat)
```

On average people lost 17 bp after the treatment.

### Subsetting data so that only a those with a positive change show up

```{r}
isPositiveChange <- which(postTreat < preTreat)

preTreatSub <- preTreat[names(isPositiveChange)]
postTreatSub <- postTreat[names(isPositiveChange)]

changeTreatSub <- preTreatSub - postTreatSub

avgChangeSub <- mean(changeTreatSub)
```

People who were positively impacted by the treatment has their bp decreased by around 20.6470 bp.

## Task 2

### Dataframe creation and creating new column for low bp

```{r}
bpTrtDf <- data.frame(names(preTreat), preTreat, postTreat, changeTreat, row.names = NULL)
names(bpTrtDf) <- c("patient", "pre_bp", "post_bp", "diff_bp")

negbp <- subset(bpTrtDf, diff_bp < 0)

# Low bp
bpTrtDf$low_bp <- bpTrtDf$post_bp < 120

# Knit
knitr::kable(bpTrtDf)
```

## Task 3

### Setting up the data also making similar to Task 2

```{r}
# Vectors
prePlacebo <- c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130)
postPlacebo <- c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124)

names(prePlacebo) <- paste("Subject", 1:10, sep = "_")
names(postPlacebo) <- paste("Subject", 1:10, sep = "_")

# Diff
changePlacebo <- prePlacebo - postPlacebo

# Making df and adding col
bpPlaceboDf <- data.frame(names(prePlacebo), prePlacebo, postPlacebo, changePlacebo)
names(bpPlaceboDf) <- c("patient", "pre_bp", "post_bp", "diff_bp")


negPlbp <- subset(bpPlaceboDf, diff_bp < 0)

# Low bp
bpPlaceboDf$low_bp <- bpPlaceboDf$post_bp < 120

```

### Making a list of all bp df and accessing the first list element (3 ways)

```{r}
bpList <- list(bpTrtDf, bpPlaceboDf)
names(bpList) <- c("treatment", "placebo")

bpList[1]
bpList$treatment
bpList[[1]]
```

### pre_bp placebo column

```{r}
bpList[[2]]$pre_bp
```

## Task 4

### Adding status col, based on bp in post trt. Using conditional logic. Done for all df in list

```{r}
bpList[[1]]$status <- character(20)
bpList[[2]]$status <- character(10)

# Loop and logic for bp trt
for (i in seq_along(bpList[[1]]$post_bp)) {
  pbpVal <- bpList[[1]]$post_bp[i]
  if (pbpVal <= 120) { 
    bpList[[1]]$status[i] <- "optimal"
  } else if (pbpVal > 130){
    bpList[[1]]$status[i] <- "high"
  } else if ((pbpVal > 120) & (pbpVal <= 130)) {
    bpList[[1]]$status[i] <-  "borderline"
  } else {
    bpList[[1]]$status[i] <- "error"
  }
}

# Loop and logic for bp placebo
for (i in seq_along(bpList[[2]]$post_bp)) {
  pbpVal <- bpList[[2]]$post_bp[i]
  if (pbpVal <= 120) { 
    bpList[[2]]$status[i] <- "optimal"
  } else if (pbpVal > 130){
    bpList[[2]]$status[i] <- "high"
  } else if ((pbpVal > 120) & (pbpVal <= 130)) {
    bpList[[2]]$status[i] <-  "borderline"
  } else {
    bpList[[2]]$status[i] <- "error"
  }
}

```
