---
title: "BradyHW4"
format: html
editor: visual
---

# Homework 4 ST 558 Group 2

```{r}
#| include: false
library(tidyverse)
library(readxl)
library(DBI)
library(RSQLite)
```

## Task 1: Conceptual Questions

Rewriting the nested call with the pipe

```{r}
# Same call written left to right with the pipe
iris |>
  as_tibble() |>
  select(starts_with("Petal"), Species) |>
  filter(Petal.Length < 1.55) |>
  arrange(Species)
```

Long format data has one observation per row and one variable per column. Wide format can pack more than one observation into a single row. We prefer long format since a lot of the functions need long format.

## Task 2: dplyr Practice

Reading in glass data and making the type a labeled factor

```{r}
# Reading in glass data (same as HW3)
glass <- read_delim("https://www4.stat.ncsu.edu/~online/datasets/glass.data",
                    delim = ",",
                    col_names = c("Id", "RI", "Na", "Mg", "Al", "Si",
                                  "K", "Ca", "Ba", "Fe", "Type_of_glass"))

# Make type a factor with real labels, then keep low iron tableware and headlamps
glass |>
  mutate(Type_of_glass = factor(Type_of_glass,
                                levels = 1:7,
                                labels = c("building_windows_float_processed",
                                           "building_windows_non_float_processed",
                                           "vehicle_windows_float_processed",
                                           "vehicle_windows_non_float_processed",
                                           "containers",
                                           "tableware",
                                           "headlamps"))) |>
  filter(Fe < 0.2, Type_of_glass %in% c("tableware", "headlamps"))
```

Reading in yeast data, dropping the columns, and summarizing by class

```{r}
# Reading in yeast data (same as HW3)
yeast <- read_table("https://www4.stat.ncsu.edu/~online/datasets/yeast.data",
                    col_names = c("seq_name", "mcg", "gvh", "alm", "mit",
                                  "erl", "pox", "vac", "nuc", "class"))

# Drop seq_name and nuc, then add mean and median of each numeric column by class
yeast |>
  select(-seq_name, -nuc) |>
  group_by(class) |>
  mutate(across(c(mcg, gvh, alm, mit, erl, pox, vac),
                list(mean = mean, median = median),
                .names = "{.col}_{.fn}"))
```

Recreating the combined wine data from HW3

```{r}
# White wine from excel, renamed using sheet 2, with a type column
white_wine <- read_excel("white-wine.xlsx")
white_names <- read_excel("white-wine.xlsx", sheet = 2)
colnames(white_wine) <- white_names$Variables
white_wine$type <- "white"

# Red wine from URL, renamed the same way, with a type column
red_wine <- read_delim("https://www4.stat.ncsu.edu/~online/datasets/red-wine.csv",
                       delim = ";")
colnames(red_wine) <- white_names$Variables
red_wine$type <- "red"

# Combine both into one data set
wine <- bind_rows(white_wine, red_wine)
```

Chaining changes to look at high quality wines

```{r}
# Filter high quality wines, sort, select columns, add alcohol mean and sd by quality
wine |>
  filter(quality > 6.5, alcohol < 132) |>
  arrange(desc(quality)) |>
  select(contains("acid"), alcohol, type, quality) |>
  group_by(quality) |>
  mutate(alcohol_mean = mean(alcohol),
         alcohol_sd = sd(alcohol))
```

## Task 3: Database Practice

Connecting to the Lahman database and listing the tables

```{r}
# Connect to the Lahman.db file in this folder
con <- dbConnect(RSQLite::SQLite(), "lahman.db")

# List all the tables
dbListTables(con)
```

Returning the Teams table for 2015 with dplyr

```{r}
# Teams table filtered to 2015
tbl(con, "Teams") |>
  filter(yearID == 2015) |>
  collect()
```

Doing the same query in SQL

```{r}
# Same 2015 Teams query as raw SQL
tbl(con, sql(
  "SELECT *
   FROM `Teams`
   WHERE (`yearID` = 2015.0)")) |>
  collect()
```

Returning all inducted hall of fame players

```{r}
# Keep inducted players and grab three columns
hof_players <- tbl(con, "HallOfFame") |>
  filter(inducted == "Y", category == "Player") |>
  select(playerID, yearID, category) |>
  collect()

hof_players
```

Adding player names from the People table

```{r}
# Pull player names
people_names <- tbl(con, "People") |>
  select(playerID, nameFirst, nameLast) |>
  collect()

# Join names onto the hall of fame players
hof_named <- inner_join(hof_players, people_names, by = "playerID")
hof_named
```

Getting each manager's win/loss record

```{r}
# Total games, wins, losses per manager, then win percentage sorted high to low
manager_record <- tbl(con, "Managers") |>
  select(playerID, G, W, L) |>
  group_by(playerID) |>
  summarize(G_managed = sum(G, na.rm = TRUE),
            Total_W = sum(W, na.rm = TRUE),
            Total_L = sum(L, na.rm = TRUE)) |>
  collect() |>
  mutate(Win_Pct = Total_W / G_managed) |>
  arrange(desc(Win_Pct))

manager_record
```

Getting people who both managed and made the hall of fame

```{r}
# All inducted hall of famers with their names
hof_all <- tbl(con, "HallOfFame") |>
  filter(inducted == "Y") |>
  select(playerID, yearID, category) |>
  collect() |>
  inner_join(people_names, by = "playerID")

# Keep only the ones who also managed
hof_managers <- hof_all |>
  inner_join(manager_record, by = "playerID")

hof_managers

# Disconnect when done
dbDisconnect(con)
```
