---
title: "Mason HW4"
format: html
editor: visual
---

```{r}
#| include: FALSE
library(tidyverse)
library(dplyr)
library(readxl)
library(DBI)
library(RSQLite)
```

## Task 1: Conceptual Questions

### Question 1:

```{r}
iris |>
  as_tibble() |>
  select(starts_with("Petal"), Species) |>
  filter(Petal.Length < 1.55) |>
  arrange(Species)
```


### Question 2:
Long format data is when each row consists of one observation and each column one variable, whereas wide format data may have more than one observation per row. We prefer long format data as most of the functions we use expect long format data.

## Task 2

Reading in glass data using code from Homework 3
```{r}
#Reading in the data
glass_data <- read_csv("https://www4.stat.ncsu.edu/online/datasets/glass.data", col_names = F, show_col_types = F)

#Setting variable names based on information from the homework
names(glass_data) <- c("Id", "RI", "Na", "Mg", "Al", "Si", "K", "Ca",
                       "Ba", "Fe", "Type_of_Glass")

#Printing tibble
glass_data
```

Overwriting Type of glass variable using `mutate()`
```{r}
glass_data |>
  mutate(Type_of_Glass = factor(Type_of_Glass, levels = 1:7,
                                labels = c("building_windows_float_processed", "building_windows_non_float_processed", "vehicle_windows_float_processed", "vehicle_windows_non_float_processed", "containers", "tableware", "headlamps"))) |>
  filter(Fe < .2, Type_of_Glass %in% c("tableware", "headlamps"))
```
