---
title: "Mason HW4"
format: html
editor: visual
---

```{r}
#| include: FALSE
library(tidyverse)
library(dplyr)
library(readxl)
library(DBI)
library(RSQLite)
```

## Task 1: Conceptual Questions

### Question 1:

```{r}
iris |>
  as_tibble() |>
  select(starts_with("Petal"), Species) |>
  filter(Petal.Length < 1.55) |>
  arrange(Species)
```


### Question 2:
Long format data is when each row consists of one observation and each column one variable, whereas wide format data may have more than one observation per row. We prefer long format data as most of the functions we use expect long format data.

## Task 2

Reading in glass data using code from Homework 3
```{r}
#Reading in the data
glass_data <- read_csv("https://www4.stat.ncsu.edu/online/datasets/glass.data", col_names = F, show_col_types = F)

#Setting variable names based on information from the homework
names(glass_data) <- c("Id", "RI", "Na", "Mg", "Al", "Si", "K", "Ca",
                       "Ba", "Fe", "Type_of_Glass")
```

Overwriting Type of glass variable using `mutate()`
```{r}
glass_data |>
  mutate(Type_of_Glass = factor(Type_of_Glass, levels = 1:7,
                                labels = c("building_windows_float_processed", "building_windows_non_float_processed", "vehicle_windows_float_processed", "vehicle_windows_non_float_processed", "containers", "tableware", "headlamps"))) |>
  filter(Fe < .2, Type_of_Glass %in% c("tableware", "headlamps"))
```

Reading in yeast data using code from Homework 3
```{r}
#Reading in the data
yeast_data <- read_table("https://www4.stat.ncsu.edu/online/datasets/yeast.data", col_names = F, show_col_types = F)

#Setting variable names based on information from the homework
names(yeast_data) <- c("Seq_name", "MCG", "GVH", "ALM", "MIT", "ERL",
                       "POX", "VAC", "NUC", "Class")
```

Starting chain to remove seq_name and nuc columns and adding mean and median columns of each numeric variable at each class grouping
```{r}
yeast_data |>
  select(-Seq_name, -NUC) |>
  group_by(Class) |>
  mutate(across(c(MCG, GVH, ALM, MIT, ERL, POX, VAC),
         list(mean = mean, median = median),
         .names = "{.col}_{.fn}"))
```

Reading in wine data using code from Homework 3
```{r}
white_data <- read_excel("white-wine.xlsx", sheet = excel_sheets("white-wine.xlsx")[1])

white_names <- read_excel("white-wine.xlsx", sheet = excel_sheets("white-wine.xlsx")[2])

colnames(white_data) <- white_names[[1]]

white_data <- white_data %>% mutate(wine_type = "white")

red_data <- read_delim("red-wine.csv", delim = ";", col_names = T,
                     show_col_types = F)

colnames(red_data) <- white_names[[1]]

red_data$wine_type = "red"

combined_wines <- bind_rows(white_data, red_data)
```

Starting chain to complete tasks given in homework
```{r}
combined_wines |>
  filter(quality > 6.5, alcohol < 13.2) |>
  arrange(desc(quality)) |>
  select(contains("acid"), alcohol, wine_type, quality) |>
  group_by(quality) |>
  mutate(mean_alcohol = mean(alcohol), sd_alcohol = sd(alcohol))
```

## Task 3: Database Practice

Downloading Lahman.db file
```{r}
con <- dbConnect(SQLite(), "lahman.db")
dbListTables(con)
```

Reading in 2015 data using dplyr
```{r}
tbl(con, "Teams") |>
  filter(yearID == 2015) |>
  collect()
```

Reading in 2015 data using SQL
```{r}
tbl(con, sql(
  "SELECT *
    FROM `Teams`
    WHERE (`yearID` = 2015.0)")
  ) |>
  collect()
```

Returning all players in the hall of fame, the year they were voted in, and their category
```{r}
hof <- tbl(con, "HallOfFame") |>
  filter(inducted == "Y", category == "Player") |>
  select(playerID, yearID, category) |>
  collect()

hof
```

Combining table from above with the People table to add the nameFirst and nameLast values
```{r}
ppl <- tbl(con, "People") |>
  select(playerID, nameFirst, nameLast) |>
  collect()

left_join(hof, ppl, by = "playerID")
```

Returning managerID, G, W, and L columns from the Managers table
```{r}
managers <- tbl(con, "Managers") |>
  select(playerID, G, W, L) |>
  group_by(playerID) |>
  summarize(G_managed = sum(G, na.rm = T),
            Total_W = sum(W, na.rm = T),
            Total_L = sum(L, na.rm = T)) |>
  collect() |>
  mutate(win_rate = Total_W / G_managed) |>
  arrange(desc(win_rate))

managers
```

Information for people that managed a team and were inducted into the hall of fame
```{r}
hof_all <- tbl(con, "HallOfFame") |>
  filter(inducted == "Y") |>
  select(playerID, yearID, category) |>
  collect() |>
  inner_join(ppl, by = "playerID")

hof_manager <- hof_all |>
  inner_join(managers, by = "playerID")

hof_manager

dbDisconnect(con)
```