---
title: "hw4"
format: html
---

## Task 1 
```{r}
# rewrite the codes 
library(tidyverse)
as_tibble(iris) |>
select(starts_with("Petal"),Species) |>
  filter(Petal.Length <
           1.55) |>
  arrange(Species)
```
What is meant by long format data and wide format data? Which do we generally prefer for statistical
analysis?
long format data refers to use one individual could have multiple observations with characteristics/features of each observation as a column, and the cell is the value of the feature. However, Wide format data combines different times of observations together as a row, so one individual only have one row with repeated observations. 
Statistical analysis prefer long format data. 

## Task 2 
```{r}
# reread that data that also names the columns. 
library(tidyverse)
glass_tbl <- read_delim("https://www4.stat.ncsu.edu/~online/datasets/glass.data", delim = ",", col_names = c("Id","RI","Na","Mg","Al","Si","K","Ca","Ba","Fe","Type of Class"))

#Start a chain on your data object that overwrites the Type_of_glass variable using mutate().
# Create a factor version of the variable that replaces the numbers with string labels
  
  glass_tbl <- glass_tbl |>
  mutate (factor_Type_of_class = case_when(
`Type of Class` == 1  ~ "building_windows_float_processed",
`Type of Class` == 2 ~ "building_windows_non_float_processed",
`Type of Class` == 3 ~ "vehicle_windows_float_processed",
`Type of Class` == 4 ~ "vehicle_windows_non_float_processed(none in this database)", 
`Type of Class` == 5 ~ "containers", 
`Type of Class` == 6 ~ "tableware", 
`Type of Class` == 7 ~  "headlamps"
  )) |>
   mutate(factor(factor_Type_of_class))
  
  #Continue your chain and keep only observations where the Fe variable is less than 0.2 and the Type of Glass is either “tableware” or “headlamp”.
  glass_tbl <- glass_tbl |>
    filter(Fe < 0.2, `Type of Class` == "tableware"|`Type of Class` == "headlamp")
  

```

YEARS data
```{r}
  #Start a chain that removes the seq_name and nuc columns.
library(tidyverse)
yeast_tbl <- read_delim("https://www4.stat.ncsu.edu/~online/datasets/yeast.data", delim = " ", col_names = c("seq_name","mcg","gvh","alm","mit","erl","pox","vac","nuc","class"))
yeast_tbl

yeast_tbl <- yeast_tbl |>
  select ("mcg","gvh","alm","mit","erl","pox","vac","class")

 # Continue your chain to add columns corresponding to the mean and median of each numeric variable


yeast_tbl <- yeast_tbl |>
  group_by(class) |>
  summarise (mean_mcg = mean(mcg),
        mean_gvh = mean(gvh),
        mean_alm = mean(alm),
        mean_mit = mean(mit),
        mean_erl = mean(erl),
        mean_pox = mean(pox),
        mean_vac = mean(vac),
        median_mcg = median(mcg),
        median_gvh = median(gvh),
        median_alm = median(alm),
        median_mit = median(mit),
        median_erl = median(erl),
        median_pox = median(pox),
        median_vac = median(vac),
        .groups = "drop")

library(tidyverse)
yeast_tbl <- read_delim("https://www4.stat.ncsu.edu/~online/datasets/yeast.data", delim = " ", col_names = c("seq_name","mcg","gvh","alm","mit","erl","pox","vac","nuc","class"))
yeast_tbl

yeast_tbl <- yeast_tbl |>
  select ("mcg","gvh","alm","mit","erl","pox","vac","class")

yeast_tbl <- yeast_tbl |>
  group_by(class) |>
  summarise (
    across(mcg:vac, 
           list(mean = mean, median = median)))

```

Wine data
```{r}
# only look at high-quality wines (quality > 6.5) and wines that have a reasonable alcohol value (alcohol < 132).
#  now sort the data from highest quality to lowest.
# select only the variables that contain acid, the alcohol variable, the type variable, and the quality variable.
# add the mean and standard deviation of the alcohol variable to the data set for each setting of the quality variable.

# read data
library(readxl)
whitewine <- read_excel("white_wine.xlsx",
                        sheet = excel_sheets("white_wine.xlsx")[1])
whitewine

# read 2st sheet
whitewine_2 <- read_excel("white_wine.xlsx",
                        sheet = excel_sheets("white_wine.xlsx")[2])
colnames(whitewine) <- whitewine_2$Variables[1:13]
whitewine

# add type column and assign "white" as type 
whitewine$type <- "white"

# read another excel 
library(readr)
redwine <- read_delim("https://www4.stat.ncsu.edu/~online/datasets/red-wine.csv", delim = ";")
colnames(redwine) <- whitewine_2$Variables[1:13]
redwine$type <- "red"

# combine two excels
allwine <- dplyr::bind_rows(whitewine, redwine)


#only look at high-quality wines (quality > 6.5) and wines that have a reasonable alcohol value (alcohol < 132).

allwine <- allwine |>
  filter (quality > 6.5, alcohol < 132)

#  now sort the data from highest quality to lowest.
allwine <- allwine |>
  arrange(desc(quality))

# select only the variables that contain acid, the alcohol variable, the type variable, and the quality variable.

allwine <- allwine |>
  select(contains("aciud"), alcohol, type, quality)

# add the mean and standard deviation of the alcohol variable to the data set for each setting of the quality variable.

allwine <- allwine |>
  group_by(quality) |>
  mutate(mean_alcohol = mean(alcohol), sd_alcohol = sd(alcohol)) |>
  ungroup()
# should we add na.rm?




```



## Task 3 
```{r}


