---
title: "LoganHW4"
format: html
---

```{r}
#| include = false
library(tidyverse)
library(readxl)
library(DBI)
library(RSQLite)
```

## Task 1

### Chaining with Base R

```{r}
arrange(filter(select(as_tibble(iris), starts_with("Petal"), Species), Petal.Length < 1.55), Species)
```

```{r}
iris |> as_tibble() |> select(starts_with("Petal"), Species) |> filter(Petal.Length < 1.55) |> arrange(Species)

```

### Question 2

Long formatted data is when data is stored in rows, wide format data stores the data in columns. Long has a row per obs and wide has more than one obs in a row. We usually prefer long the make operations in code easier.

## Task 2

### Glass data

```{r}
glassData <- read_csv("https://www4.stat.ncsu.edu/online/datasets/glass.data", 
                      col_names = c("Id", "RI", "Na", "Mg", "Al", "Si", "K", "Ca", "Ba", "Fe", "Type_of_glass"),
                      show_col_types = FALSE)

# Overwriting Type_of_Glass
glassData %>% mutate(Type_of_glass = factor(Type_of_glass,
                                levels = c(1, 2, 3, 4, 5, 6, 7),
                                labels = c("building_windows_float_processed",
                                           "building_windows_non_float_processed",
                                           "vehicle_windows_float_processed",
                                           "vehicle_windows_non_float_processed",
                                           "containers",
                                           "tableware",
                                           "headlamps"))) %>% 
              filter(Fe < 0.2, Type_of_glass == "tableware" | Type_of_glass == "headlamps")

```

### Yeast data

```{r}
yeastData <- read_table("https://www4.stat.ncsu.edu/online/datasets/yeast.data",
                        col_names = c("seq_name", "mcg", "gvh", "alm", "mit", "erl" , "pox", "vac", "nuc", "class"),
                        show_col_types = FALSE) 

yeastData %>% select(-c("seq_name", "nuc")) %>%
  group_by(class) %>%
  mutate(across(c(mcg, gvh, alm, mit, erl, pox, vac), list(mean = mean, median = median),.names= "{.col}_{.fn}"))
```

### Wine data

```{r}
# Reading in the data for white
wineWhiteData <- read_excel("white-wine.xlsx")

# Reading in names for the data from sheet 2
wineWhiteDataNames <- read_excel("white-wine.xlsx",
  sheet = excel_sheets("white-wine.xlsx")[2])

# Renaming
colnames(wineWhiteData) <- wineWhiteDataNames[[1]]

# Adding col to indicate wine type
wineWhiteData <- wineWhiteData %>% mutate(type = "white")

# Reading in the data for red
wineRedData <- read_delim("red-wine.csv",
                          delim = ";",
                          col_names = TRUE,
                          show_col_types = FALSE)

# Renaming, has same names as white
colnames(wineRedData) <- wineWhiteDataNames[[1]]

# Adding col to indicate wine type also making sure everything is double for later operations
wineRedData <- wineRedData %>% mutate(across(wineWhiteDataNames[[1]], as.double)) %>% 
  mutate(type = "red")

# Combining
wineData <- bind_rows(wineWhiteData, wineRedData)

# Look at high quality wines
wineData <- wineData %>% filter(quality > 6.5, alcohol < 13.2) %>%
  arrange(desc(quality)) %>% select(contains("acid"), alcohol, type, quality) %>%
  group_by(quality) %>% mutate(mean_alcohol = mean(alcohol), sd_alcohol = sd(alcohol))

```

## Task 3

### Connecting to the Lahman database

```{r}
con <- dbConnect(RSQLite::SQLite(), "lahman.db")

# Look at all tables
dbListTables(con)

# Returning Teams table for 2015 using dplyr
tbl(con, "Teams") %>%
  filter(yearID == 2015) %>% collect()

# Returning Teams table for 2015 using SQL
tbl(con, sql(
  "SELECT *
  FROM `Teams`
  WHERE (`yearID` = 2015.0)")
) %>% collect()
```

### Queries

#### Returning all players in hall of fame

```{r}
# Only want player, year, and category vars
hofPlayers <- tbl(con, "HallOfFame") %>% 
  filter(inducted == "Y", category == "Player") %>%
  select(playerID, yearID, category) %>% collect()
hofPlayers
```

#### Combining previous with People
```{r}
# Storing people db
ppl <- tbl(con, "People") %>% select(playerID, nameFirst, nameLast) %>% collect()

# Joining
left_join(hofPlayers,ppl, by = "playerID")
```

#### Manager's wr

```{r}
managerWr <- tbl(con, "Managers") %>%
  select(playerID, G, W, L) %>%
  group_by(playerID) %>%
  summarize(G_managed = sum(G, na.rm = TRUE),
    Total_W = sum(W, na.rm = TRUE),
    Total_L = sum(L, na.rm = TRUE)) %>%
  collect() %>%
  mutate(win_rate = (Total_W / G_managed)) %>%
  arrange(desc(win_rate))

managerWr
```

#### Information from people that managed a team inducted into hof

```{r}
# HOF table 
hofAll <- tbl(con, "HallOfFame") %>%
  filter(inducted == "Y") %>%
  select(playerID, yearID, category) %>% collect() %>% left_join(ppl, by = "playerID")

hofManagerWr <- hofAll %>%
  inner_join(managerWr, by = "playerID")

hofManagerWr

# Disconnect
dbDisconnect(con)
```


