---
title: "Ryan's HW4"
format: html
---

## Task 1: Conceptual Questions

1. 
```{r}
library(tidyverse)
iris |>
  as_tibble() |>
  select(starts_with("Petal"), Species) |>
  filter(Petal.Length < 1.55) |>
  arrange(Species)
```

2. Long format data is when each row is a single observation and each column is a single variable. Wide format data is when a single row can contain multiple observations. We tend to prefer long format because most plotting, summarizing, and modeling functions expect long format.

## Task 2:

For this task we manipulate the data sets we read in for HW 3, starting with the glass data set.
```{r}
library(tidyverse)

# Read in glass data
glass_data <- read_csv(
  "https://www4.stat.ncsu.edu/~online/datasets/glass.data",
  col_names = c("Id", "RI", "Na", "Mg", "Al", "Si", "K", "Ca", "Ba", "Fe", "Type_of_glass")
)

# Piping several changes (Type as factor with appropriate labels, subsetting Type and Fe)
glass_data <- glass_data |>
  mutate(Type_of_glass = factor(Type_of_glass,
                                levels = c(1, 2, 3, 4, 5, 6, 7),
                                labels = c("building_windows_float_processed",
                                           "building_windows_non_float_processed",
                                           "vehicle_windows_float_processed",
                                           "vehicle_windows_non_float_processed",
                                           "containers",
                                           "tableware",
                                           "headlamps"))) |>
  filter(Fe < 0.2, Type_of_glass == "tableware" | Type_of_glass == "headlamps")

# Print tibble
glass_data
```

Next we perform manipulations on the yeast data set.
```{r}
# Read in yeast data
yeast_data <- read_fwf(
  "https://www4.stat.ncsu.edu/~online/datasets/yeast.data",
  fwf_widths(
    widths = c(12, 6, 6, 6, 6, 6, 6, 6, 6, 3),               # Character widths for each column
    col_names = c("seq_name", "mcg", "gvh", "alm", "mit",
                  "erl", "pox", "vac", "nuc", "class")
  )
)

# Piping several changes (remove seq_name and nuc, print mean and median for remaining cols)
yeast_data <- yeast_data |>
  select(mcg, gvh, alm, mit, erl, pox, vac, class) |>
  group_by(class) |>
  mutate(across(c(mcg, gvh, alm, mit, erl, pox, vac), 
                list(mean=mean, median=median), 
                .names= "{.col}_{.fn}"))

# Print tibble
yeast_data
```

Lastly, we manipulate the wine data set. However first we must recreate it from the white and red wine data.
```{r}
library(readxl)
### WHITE WINE DATA ###
# Read in white wine data
white_wine <- read_excel("white-wine.xlsx")

# Specifying the second sheet
white_wine_names <- read_excel("white-wine.xlsx", sheet = 2)

# Renaming columns
colnames(white_wine) <- white_wine_names$Variables

# Column add
white_wine$type <- "white"

### RED WINE DATA ###
# Read in red wine data
red_wine <- read_delim("red-wine.csv", delim = ";")

# Renaming columns
colnames(red_wine) <- white_wine_names$Variables

# Column add
red_wine$type <- "red"

### WINE DATA ###
# Combine data sets
wine_data <- dplyr::bind_rows(white_wine, red_wine)

# Print tibble
wine_data
```

Now that we have the combined wine data set, we can manipulate it.
```{r}
# Piping several changes to the wine data set
wine_data <- wine_data |>
  filter(quality > 6.5, alcohol < 132) |>
  arrange(desc(quality)) |>
  select(fixed_acidity, volatile_acidity, citric_acid, alcohol, type, quality) |>
  group_by(quality) |>
  mutate(across(alcohol, 
                list(mean = mean, sd = sd), 
                .names = "{.col}_{.fn}"))

# Print tibble
wine_data
```

## Task 3: Database Practice

Here we connect to the Lahman.db file.
```{r}
library(DBI)

# Connecting to Lahman.db
con <- dbConnect(RSQLite::SQLite(), "lahman.db")
dbListTables(con)
```

With Lahman.db now loaded, we return all of the data from the Teams table for the year 2015. 
```{r}
library(dplyr)

tbl(con, "Teams") |>
  filter(yearID == 2015) |>
  collect() # Returning without lazy evaluation
```

We do the same thing but now with SQL
```{r}
tbl(con, sql(
"SELECT `Teams`.*
FROM `Teams`
WHERE (`yearID` = 2015.0)")
)|>
  collect() # Returning without lazy evaluation
```

Here we return specific players from the Hall of Fame.
```{r}
last_HOF <- tbl(con, "HallOfFame") |>
  filter(category == "Player", inducted == "Y") |> # Inducted (not just nominated)
  select(playerID, yearID, category) |>
  collect() # Returning without lazy evaluation

# Print tibble
last_HOF
```

Next we merge our manipulated HallOfFame data with the People table.
```{r}
# Retrieving names and IDs from the People data
first_HOF <- tbl(con, "People") |>
  select(playerID, nameFirst, nameLast) |>
  collect()

# Print tibble
first_HOF

# Join data sets
first_last_HOF <- inner_join(first_HOF, last_HOF, by = "playerID") |>
  collect()

# Print tibble
first_last_HOF
```

Moving on, we examine the Managers table.
```{r}
tbl(con, "Managers") |>
  select(playerID, G, W, L) |>
  group_by(playerID) |> # Technically the Manager ID
  summarize(G_managed = sum(G, na.rm = TRUE),
            Total_W = sum(W, na.rm = TRUE),
            Total_L = sum(L, na.rm = TRUE)) |>
  collect() |>
  mutate(Win_Pct = Total_W / G_managed) |>
  arrange(desc(Win_Pct))
```

Lastly, we will build off of our work from the previous data sets to return info about the
people that managed a team and were inducted into the hall of fame.
```{r}
# HOF table w/o restricting "Player" category
HOF_info <- tbl(con, "HallOfFame") |>
  filter(inducted == "Y") |>
  select(playerID, yearID, category) |>
  inner_join(tbl(con, "People") |> 
               select(playerID, nameFirst, nameLast),
             by = "playerID")


# Managers table (same as Q16)
managers_info <- tbl(con, "Managers") |>
  select(playerID, G, W, L) |>
  group_by(playerID) |>
  summarize(G_managed = sum(G, na.rm = TRUE),
            Total_W = sum(W, na.rm = TRUE),
            Total_L = sum(L, na.rm = TRUE))


# Combine data sets - people who are BOTH inducted AND are managers
HOF_managers <- HOF_info |>
  inner_join(managers_info, by = "playerID") |>
  collect() |>                                    # Moving this part down from the managers code
  mutate(Win_Pct = Total_W / G_managed) |>
  arrange(desc(Win_Pct))

HOF_managers

dbDisconnect(con)
```

